---
title: "Workflow basics"
author: "Fabio Favoretto"
date: "`r Sys.Date()`"
output: 
        ioslides_presentation:
                logo: "figs/cropped-rstudio.png"
                widescreen: true
                incremental: true
---

## References

- https://r4ds.had.co.nz/workflow-basics.html
- https://r4ds.had.co.nz/workflow-scripts.html


# Coding basics 
```{r, include=FALSE}
knitr::opts_chunk$set(eval=FALSE, warning = FALSE, message = FALSE)
```


## You can use R as a calculator: 
```{r}
1 / 200 * 30

(59 + 73 + 2) / 3

sin(pi / 2)

```

You can create new objects with the symbol `<-`
Press `alt` + `-` in Windows as a shortcut

```{r}
x <- 3 * 4

object_name <- value

```


## Importance in names: 

i_use_snake_case

otherPeopleUseCamelCase

some.people.use.perionds

And_aFew.People_RENOUNCEconvetion

use TAB to autocomplete the name


## let's make another one
```{r}
r_rocks <- 2 ^ 3

r_rock
R_rocks

my_variable <- 10
my_variable -5

```


## More stuff
There’s an implied contract between you and R: it will do the tedious computation for you, but in return, you must be completely precise in your instructions. 

Typos matter. Case matters.

## Calling functions 

function_name(arg1 = val1, arg2 = val2, ...)

we try using seq() which makes regular sequences of numbers

try also using TAB completion

```{r}
seq()
seq(1, 10)
#seq(1, 10

    #R wants to help you if you miss something.

y <- seq(1, 10, length.out = 5)

```

try using TAB completions inside a function

## Similarly, you can use the `c()` function

```{r}
x <- c(1:8, 75, 34)
x/2
```


# Practice

## Why the code below does not work?

```{r}
my_variable <- 10
my_varıable

```


## Why the code below does not work?

```{r}
x <- (2, 4, 5, 6, 7)

```


## Why the code below does not work?

```{r}
myname <- Fabio
```



# Scripting

## try this in R
```{r}
library(dplyr)
library(nycflights13)

not_cancelled <- flights %>% 
        filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
        group_by(year, month, day) %>% 
        summarise(mean = mean(dep_delay))
```

