---
title: "R-functions and iterations"
author: "Fabio Favoretto"
date: "`r Sys.Date()`"
output: 
        ioslides_presentation:
                logo: "figs/cropped-rstudio.png"
                widescreen: true
                incremental: true
---



# R-Functions
```{r, include=FALSE}
knitr::opts_chunk$set(eval=FALSE, warning = FALSE, message = FALSE)
```

## What is a function?

Basically resume in one word several task R should do and can be applied to different datasets.

You should consider writing a function whenever you’ve copied and pasted a block of code more than twice.


## An example

we can write a function that converts Fahrenheit to Kelvin 
```{r}
fahr_to_kelvin <- function(temp) {
  kelvin <- ((temp - 32) * (5 / 9)) + 273.15
  return(kelvin)
}

```

## Excercise

Try writing a function to calculate kelvin to celsius

## Solution

```{r}
kelvin_to_celsius <- function(temp) {
  celsius <- temp - 273.15
  return(celsius)
}
```

## Defensive programming

You can specify several things and control programming to check conditions to operate your function:

```{r}
fahr_to_kelvin <- function(temp) {
  stopifnot(is.numeric(temp))
  kelvin <- ((temp - 32) * (5 / 9)) + 273.15
  return(kelvin)
}
```

## Excercise 

Write a function to convert farheneit to celsius using "defensive programming"

```{r}
fahr_to_celsius <- function(temp) {
  stopifnot(is.numeric(temp))
  celsius <- (temp - 32) / 1.8 
  return(celsius)
}
```


## Another example

We want a function that quickly calculate the mean of a variable in a defined region: 

```{r eval=FALSE}

library(readxl) # this is to load the excel
library(tidyverse) # this is for everything else...


# Loading data ------------------------------------------------------------

df <- read_xlsx("data/fishing_effort.xlsx") # this loads data

meansum <- function(Region, Site, variable=Pangas, group = "LaPaz") {
  group_by(Region, Site) %>%
                  filter(Region == group) %>%
                  summarise(mean = mean(Pangas))
}

meansum(df, group = "CaboPulmo", variable=PFE)

```

## Iterations

```{r eval=FALSE}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

```{r eval=FALSE}

map(df, mean)
map(df, median)
map(df, sd)

```


## Types of `map`

`map()` makes a list.
`map_lgl()` makes a logical vector.
`map_int()` makes an integer vector.
`map_dbl()` makes a double vector.
`map_chr()` makes a character vector.


## Models

```{r eval=FALSE}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))

models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))

models %>% 
  map(summary) %>% 
  map_dbl(~.$r.squared)

models %>% 
  map(summary) %>% 
  map_dbl("r.squared")
```


## Exercises

Write code that uses one of the map functions to:

- Compute the mean of every column in `mtcars`.
- Determine the type of each column in `nycflights13::flights`.
- Compute the number of unique values in each column of `iris`.


## Iteration with large df
[purrr-cheatsheet](cheatsheets/purrr.pdf)

```{r eval=FALSE}
n_iris <- iris %>% 
        group_by(Species) %>% 
        nest()

mod_fun <- function(df)
lm(Sepal.Length ~ ., data = df)
m_iris <- n_iris %>%
mutate(model = map(data, mod_fun))

b_fun <- function(mod)
coefficients(mod)[[1]]
m_iris %>% transmute(Species,
beta = map_dbl(model, b_fun))

```


