---
title: "Markdown"
author: "Fabio Favoretto"
date: "`r Sys.Date()`"
output: 
        html_document:
          fig.captions: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Reference

https://bookdown.org/yihui/rmarkdown/


# Inline formatting

Inline text will be italic if surrounded by underscores or asterisks, e.g., _text_ or *text*. Bold text is produced using a pair of double asterisks (**text**). A pair of tildes (~) turn text to a subscript (e.g., H~3~PO~4~ renders H3PO4). A pair of carets (^) produce a superscript (e.g., Cu^2+^ renders Cu2+).

To mark text as `inline code`, use a pair of backticks, e.g., `code`. 


Hyperlinks are created using the syntax [text](link), e.g., [RStudio](https://www.rstudio.com). The syntax for images is similar: just add an exclamation mark, e.g., `![alt text or image title](path/to/image)`. Footnotes are put inside the square brackets after a caret ^[], e.g., ^[This is a footnote.].



Section headers can be written after a number of pound signs, e.g.,

# First-level header

## Second-level header

### Third-level header
If you do not want a certain heading to be numbered, you can add {-} or {.unnumbered} after the heading, e.g.,

# Preface {-}
Unordered list items start with *, -, or +, and you can nest one list within another list by indenting the sub-list, e.g.,

- one item
- one item
- one item
    - one more item
    - one more item
    - one more item
    

Ordered list items start with numbers (you can also nest lists within lists), e.g.,

1. the first item
2. the second item
3. the third item
    - one unordered item
    - one unordered item
    

Blockquotes are written after >, e.g.,

> "I thoroughly disapprove of duels. If a man should challenge me,
  I would take him kindly and forgivingly by the hand and lead him
  to a quiet place and kill him."
>
> --- Mark Twain


In R, the character
# indicates a comment.
And does “-” mean a bullet point below?

The result of 5
- 3 is 2.
Different flavors of Markdown may produce different results if there are no blank lines.



# R code chunks and inline R code

You can insert an R code chunk either using the RStudio toolbar (the Insert button) or the keyboard shortcut Ctrl + Alt + I (Cmd + Option + I on macOS).

There are a lot of things you can do in a code chunk: you can produce text output, tables, or graphics. You have fine control over all these output via chunk options, which can be provided inside the curly braces (between ```{r and }). For example, you can choose hide text output via the chunk option results = 'hide', or set the figure height to 4 inches via fig.height = 4. Chunk options are separated by commas, e.g.,

```{r, fig.align='center', fig.cap="Figure 1: this is a caption for the figure", message=FALSE, warning=FALSE}

library(tidyverse)
ggplot(iris,
       aes(
         x = Sepal.Length,
         y = Sepal.Width,
         col = Species,
         shape = Species
       )) +
  geom_point(size = 6, alpha = 0.6) +
  theme_classic()

```



## Why not doing a interactive graph? :) 

```{r, fig.align='center', fig.cap="Figure 2: a cool interactive graph!", message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(plotly)

# Scatterplot
p <-
  ggplot(iris,
         aes(
           x = Sepal.Length,
           y = Sepal.Width,
           color = Species,
           shape = Species
         )) +
  geom_point(size = 6, alpha = 0.6)+
  theme_classic()


ggplotly(p)
```




Now we want to go further, and do an interactive MAP!!! 
Let's use the fishing effort dataset: 

```{r echo=FALSE, fig.align='center', fig.cap="Figure 3: map of fishing effort in the GoC", message=FALSE, warning=FALSE}

df <- readxl::read_xlsx("data/fishing_effort.xlsx")

library(sf)
library(mapview)
df_sp <- st_as_sf(x = df, 
                        coords = c("Longitude", "Latitude"),
                        crs = "+proj=longlat +datum=WGS84")


mapview(df_sp, zcol="PFE", cex="PFE")

```

